Supplementary Material

### 2025-05-30 manuscript/manuscript.qmd
major update (see description =>)
- adjusted model now include fat and muscle as covariates
- hand-made and R-made figures updated to reflect the new posteriors and predictions
- adjustements made also on the main manuscript and supplementary material

**Material and Methods**: Using a cross-sectional design, 81 community-dwelling older adults (mean age 70.7 +/- 5.8 years) underwent immune profiling to quantify total B cells and four CD21/CD11c phenotypes. Continuous R-R interval (RRi) data were recorded at rest, during a Two-Minute Step Test (TMST), and over a five-minute recovery period. A coupled-logistic RRi-vs-time model yielded seven kinetic parameters per participant. Individual parameter estimates were regressed on standardized immune predictors using multivariate Bayesian models, adjusting for age and sex.

**Results**: Higher counts of CD21^+^/CD11c^+^ B cells were associated with elevated baseline RRi (resting vagal tone), a moderated exercise-induced RRi drop, and slower post-exercise recovery (posterior probability > 95%). Conversely, greater CD21^-^/CD11c^- B-cell counts related with lower resting RRi, a pronounced sympathetic-driven RRi decrease during exercise, and more rapid vagal reactivation during recovery (posterior probability > 90%).

@@

@@

@@

where $x_{i1},\dots,x_{i6}$ denote the six centered and scaled immune-cell predictors (absolute lymphocyte count, total B-cell count, and the four B-cell phenotypes by CD21/CD11c combinations). In the first model, we regressed $\widehat{\theta}_{ik}$ on those six immune predictors via @eq-second-model. In the second model, we extended @eq-second-model by adding two additional terms, $\beta_{k,\mathrm{age}} (\mathrm{age}_i)$ and $\beta_{k,\mathrm{sex}} (\mathrm{sex}_i)$, to account for confounding by age and sex when modeling RRi kinetics. Prior to fitting either multivariate model, all continuous predictors (immune counts and age) were centered to their mean and scaled by their standard deviation, and sex was coded as 0/1. We specified $\mathcal{N}(0,3)$ priors on every main-effect coefficient $\beta_{k}$ to exert a weakly regularizing influence and to mitigate the impact of outliers.

@@

@@

[**Table 1**]{#tab-1}. Sample characteristics in sociodemographical, immunological variables.

@@

data[, sex := factor(sex, levels = c("F", "M"), labels = c("Females", "Males"))]
tbl_data <- data[, c(15:16,9:14)]
tbl_data |> 
  tbl_summary(by = sex,
            missing = "no",
            statistic = list(
              all_continuous() ~ "{mean} ± {sd}"
            ),
            label = list(
              age ~ "Age (years old)",
              abc_linfocitos_total ~ "Total lymphocytes (n)",
              abc_linfocitosb_total ~ "Total B cells (n)",
              cd21_m_cd11c_p ~ "CD21- CD11c+ B cells (n)",
              cd21_p_cd11c_p ~ "CD21+ CD11c+ B cells (n)",
              cd21_p_cd11c_m ~ "CD21+ CD11c- B cells (n)",
              cd21_m_cd11c_m ~ "CD21- CD11c- B cells (n)"
            )) |> 
  add_overall() |> 
  add_difference(test = list(all_continuous() ~ "smd")) |> 
  modify_spanning_header(all_stat_cols(stat_0 = F) ~ "**Sex**")
```

@@

![](../figures/hand-designed/rri-curve.pdf){width=100%}

@@

<!--- ## See output from "R/rri-characteristics.R"
   Parameters Estimate    SE               CI
       <fctr>    <num> <num>           <char>
1:      alpha   876.89 15.29 [845.45, 906.55]
2:       beta   386.64 17.99 [350.97, 420.27]
3:          c     0.90  0.02     [0.86, 0.93]
4:     lambda     2.96  0.16     [2.63, 3.26]
5:        phi     2.29  0.20     [1.91, 2.69]
6:        tau     6.75  0.08     [6.58, 6.91]
7:      delta     2.47  0.08      [2.3, 2.62]
--->

In our sample, the baseline RRi controlled by $\alpha$ was 876.9 ms (CI~95%~[845.5, 906.6]), with an exercise induced drop ($\beta$) of 386.6 (CI~95%~[351.0, 420.3]). The recovery proportion ($c$ parameter) relative to $\beta$ was 0.9 (CI~95%~[0.86, 0.93]).

In relation to the rate parameters, controlling the speed of transition between exercise and recovery phases, we observed a rate of exercise-induced decay ($\lambda$) of 2.96 min^-1^ (CI~95%~[2.63, 3.26]) and a post exercise recovery rate ($\phi$) of 2.29 min^-1^ (CI~95%~[1.91, 2.69]).

For time parameters controlling the timing of the aforementioned RRi kinetics, we observed that the time of the exercise-induced ($\tau$) drop aligned at 6.75 (CI~95%~[6.58, 6.91]) and a duration of the exertion-related RRi depression ($\delta$) was prolonged for 2.47 (CI~95%~[2.3, 2.62]).

@@

![](../figures/hand-designed/rri-signatures.pdf){width=100%}

@@

For individuals with more expression of both CD21 and CD11c (i.e., double positive), we observed a higher RRi at rest (ES on $\alpha$ = 0.52, CI~95%~[0.18, 0.86], pd = 99.8%, ps = 98.9%), more pronounced exercise-induced RRi decrease (ES on $\beta$ = 0.4, CI~95%~[0.06, 0.75], pd = 98.8%, ps = 95.6%), less post-exercise RRi recovery (ES on $c$ = -0.28, CI~95%~[-0.61, 0.06], pd = 95.4%, ps = 85.7%) and higher duration of RRi depression till recovery kinetics begin relative to RRi drop (ES on $\delta$ = 0.3, CI~95%~[-0.08, 0.67], pd = 94.3%, ps = 85.7%).

@@

On the other side, individuals with higher counts of B cells absent of CD21 or CD11c markers (i.e., double negative) experienced lower baseline RRi (ES on $\alpha$ = -0.29, CI~95%~[-0.67, 0.1], pd = 92.9%, ps = 83.1%), less exercise-induced RRi decrease (ES on $\beta$ = -0.41, CI~95%~[-0.8, -0.03], pd = 98.2%, ps = 94.7%), better post-exercise recovery (ES on $c$ = 0.31, CI~95%~[-0.05, 0.69], pd = 94.9%, ps = 86.6%), steeper RRi drop (ES on $\lambda$ = 0.4, CI~95%~[-0.02, 0.81], pd = 97.2%, ps = 92.2%) and early initiation of cardiac autonomic kinematics (ES on $\tau$= -0.38, CI~95%~[-0.77, 0.05], pd = 96.3%, ps = 90.7%).

@@

For B cells with more CD11c expression, but not CD21, we observed that greater counts of this cell phenotype were associated 73% probability of having a significant slower RRi drop initiated by exercise (ES on $\lambda$ = -0.21, CI~95%~[-0.58, 0.15], pd = 88%, ps = 73%) and close to 90% probability of observing a significant delay in exercise-induced cardiac autonomic kinematics (ES on $\tau$ = 0.32, CI~95%~[-0.04, 0.69], pd = 96%, ps = 89%).

The specific effect of each specific B cell phenotype on the different aspects of RRi dynamics can be seen in [Figure 4](#fig-4).

![](../figures/fig_params_vs_immune_adjusted.pdf){width=100%}

[**Figure 4**](#fig-4). Posterior distribution of the standardized effect of B cell CD21 and CD11c expression on the magnitude of model parameters influencing the exercise-induced cardiac autonomic dynamics. $\alpha$ controls the baseline RRi; $\beta$ is proportional to the RRi drop, induced by exercise; $c$ is the recovery proportion, relative to $\beta$; $\lambda$ and $\phi$ are the steepness of exercise-induced RRi drop and post exercise recovery; $\tau$ correspond to the time at which the exercise-induced RRi drop occurs; $\delta$ correspond to the time-duration of the RRi depression, relative to $\tau$, that takes before the recovery kinetics begin.

@@

Our findings reveal a compelling dichotomy in the relationship between specific B cell subpopulations and autonomic dynamics during exercise in older adults. Individuals with a higher proportion of CD21^+^/CD11c^+^ (double positive) B cells exhibited characteristics of heightened parasympathetic activity at rest, including elevated baseline vagal tone, longer resting RR intervals, and greater initial heart rate variability. This group also displayed a blunted sympathetic response during exercise, indicated by a smaller abrupt drop in RRi, and a protracted cardiac recovery post-exercise. Conversely, participants with a predominance of CD21^-^/CD11c^-^ (double negative) B cells showed the opposite autonomic profile: low baseline vagal tone, characterized by shorter RR intervals and lower initial variability. They also presented an exaggerated sympathetic response to physical exertion, reflected in a marked decrease in RRi, and a more rapid return to baseline heart rate following exercise. These contrasting observations suggest that the relative abundance of these B cell subpopulations may serve as an indicator of an individual's autonomic balance, with the CD21^+^/CD11c^+^ phenotype aligning with a vagally dominant state and dampened sympathetic reactivity, while the CD21^-^/CD11c^-^ phenotype corresponds to a more sympathetically inclined state coupled with efficient compensatory vagal capacity.

@@

The attenuated sympathetic response during exercise observed in this group is another notable finding. While a high resting vagal tone typically leads to a more pronounced vagal withdrawal at the onset of exercise, our data suggest a limited initial heart rate acceleration, potentially due to cholinergic modulation of the autonomic nervous system (ANS) [@giunta2024autonomic]. Cholinergic activation has been shown to counteract sympathetic signals, and a blunted adrenergic response might also limit the exercise-induced mobilization of lymphocytes [@smith2025exercise], potentially preventing excessive immune activation in response to physical stress, thus suggesting a more controlled immune state in individuals with a dominant vagal influence.

However, this vagal-centric profile was also characterized by a slower return to baseline heart rate after exercise. Despite a less pronounced initial increase in heart rate, individuals with more CD21^+^/CD11c^+^ B cells took longer to recover. This protracted recovery could stem from a less efficient re-establishment of autonomic balance following a moderated sympathetic response, where the sympathetic nervous system (SNS) might not have been sufficiently activated to trigger a strong subsequent vagal rebound. Alternatively, it might indicate a degree of autonomic rigidity [@castillo2025enhancing], where a predominantly parasympathetic system at rest exhibits difficulty in rapidly adjusting after a perturbation. The lower $c$ parameter (recovery proportion of the RRi curve) observed in this group supports the notion of reduced autonomic flexibility, suggesting that while a basal anti-inflammatory environment might be present, the capacity for rapid adaptation to acute physiological challenges could be limited.

The characteristics associated with a higher proportion of CD21^+^/CD11c^+^ B cells find parallels in active older adults, who often display high resting HRV and moderated cardiovascular responses to exercise [@mabe2024associations], along with less inflammatory immune profiles [@tylutka2021lifestyle]. This suggests that these memory B cells might serve as both a marker and a modulator of this physiological state. While CD11c^+^ B cells (ABCs) in healthy individuals can differentiate into plasmablasts upon stimulation [@golinski2020cd11c], they also exhibit some resistance to regulatory signals like IL-10 [@golinski2020cd11c], implying potential modulation by neuroendocrine pathways. Future research into the effects of interventions such as chronic vagal nerve stimulation on these cells could be valuable in exploring ways to enhance their anti-inflammatory profile while preserving their capacity for immune response when needed.

@@

During exercise, this subpopulation exhibited an amplified autonomic response, with the marked tachycardia, reflected in a high $\beta$ parameter, indicating a robust sympathetic discharge. This heightened response is consistent with their low baseline parasympathetic tone; with less vagal "braking", the cardiovascular system responds rapidly and intensely to the exercise stimulus. While potent acute sympathetic activation has advantages such as increased cardiac output and immune cell mobilization [@smith2025exercise], potentially contributing to the clearance of senescent cells through exercise [@simpson2021exercise], it also carries potential risks, including cardiac arrhythmias and increased inflammation.

A notable and potentially beneficial finding was the excellent autonomic recovery (high $c$ parameter) observed in the CD21^-^/CD11c^-^ group despite their lower initial vagal tone. This suggests a compensatory and efficient vagal reactivation following the sympathetic surge, possibly mediated by baroreceptor and mechanoreceptor reflexes responding to the exercise-induced cardiovascular changes [@besnier2017exercise]. This efficient cholinergic recovery could aid in resolving immune activation once the exercise stimulus ceases. Thus, individuals with a more sympathicotonic profile might be capable of mounting strong but transient responses, followed by a relatively rapid return to autonomic balance mediated by the parasympathetic nervous system. This pattern resonates with descriptions of mild autonomic insufficiency, where sympathetic hyper-responsiveness can coexist with late vagal hyper-reactivity [@besnier2017exercise], and aligns with the concept of autonomic resilience [@castillo2025enhancing].

This autonomic and immune profile shares similarities with clinical conditions such as metabolic syndrome and obesity, where reduced HRV and high basal sympathetic activity are often observed, although interventions can improve vagal reactivity [@frasca2023obesity]. The link between alterations in B cells and autonomic behavior suggests a bidirectional relationship, where not only can the immune profile influence the autonomic response, but also that modulating the ANS can potentially impact age-associated immune alterations, as seen with chronic exercise interventions [@weyh2020physical; @tylutka2021lifestyle].

@@

This study offers a novel contribution by integrating autonomic physiology, through nonlinear modeling of RR-interval dynamics, with detailed B-cell phenotyping in the context of healthy aging. The application of a coupled logistic framework allowed us to capture temporally resolved autonomic shifts during exercise and recovery, providing a more dynamic assessment compared to standard HRV metrics. Our findings, which highlight distinct associations between CD21/CD11c B-cell subsets and specific autonomic parameters, offer new avenues for understanding immune-ANS interactions.

However, several limitations warrant consideration. The cross-sectional design of our study precludes the establishment of causal inferences between the observed immune and autonomic characteristics. Furthermore, while surface marker-based phenotyping allowed for the identification of specific B-cell subpopulations, it did not elucidate their functional or mechanistic relevance. The absence of functional assays or transcriptomic profiling to complement the flow cytometry data, along with the unquantified potential for batch or gating effects, represents a further limitation. Methodologically, our two-stage modeling pipeline, though computationally advantageous, did not propagate posterior uncertainty between steps, underscoring the need for future work employing fully joint hierarchical inference. Finally, despite adjusting for key demographic factors, we cannot exclude the possibility of residual confounding by variables such as fitness levels, BMI, or circadian influences.

Moving forward, future research should aim to overcome these limitations by employing longitudinal study designs to explore causality, incorporating functional and mechanistic analyses to understand the roles of the identified B-cell populations, and utilizing more integrated statistical modeling approaches to further elucidate the complex interplay between the immune and autonomic systems in aging.

@@

This research suggests that B cell subpopulations reflect and potentially modulate ANS dynamics during exercise in older adults. An integrative neuro-immunological approach provides a more comprehensive understanding of the health status of older individuals, identifying "pro-resilient" versus "pro-fragile" profiles that might not be evident when evaluating systems separately. Individuals with a predominance of CD21^+^/CD11c^+^ B cells exhibit higher baseline vagal tone, a blunted sympathetic response during exercise, and slower autonomic recovery. In contrast, those with CD21-/CD11c- B cells present lower baseline vagal tone, a more pronounced sympathetic response, and faster post-exercise recovery. These results indicate that CD21^+^/CD11c^+^ B cells are associated with a more regulated and less inflammatory autonomic profile, whereas CD21^-^/CD11c^-^ B cells reflect a state of chronic inflammation. Exercise could modulate these autonomic and immune profiles, thereby improving cardiovascular health in older adults. This investigation underscores the importance of an integrative neuro-immunological approach for developing predictive biomarkers and personalized strategies to optimize health in aging.

>>>>

**Material and Methods**: Using a cross-sectional design, 81 community-dwelling older adults (mean age 70.7 ± 5.8 years) underwent immune profiling to quantify total B cells and four CD21/CD11c phenotypes. Continuous R-R interval (RRi) data were recorded at rest, during a Two-Minute Step Test (TMST), and over a five-minute recovery period. A coupled-logistic RRi-vs-time model yielded seven kinetic parameters per participant. Individual parameter estimates were regressed on standardized immune predictors using multivariate Bayesian models, adjusting for age, sex and body composition variables.

**Results**: Higher counts of CD21^+^/CD11c^+^ B cells were associated with elevated baseline RRi (resting vagal tone), an increased exercise-induced RRi drop, and an incomplete post-exercise recovery (posterior probability > 90%). Conversely, greater CD21^-^/CD11c^- B-cell counts related with lower resting RRi, a pronounced sympathetic-driven RRi decrease during exercise, and more complete vagal reactivation during recovery (posterior probability > 90%).

@@

Following the initial registry, body composition measurements were performed using multi-frequency bioelectrical impedance analysis (BIA) according to standardized procedures. Participants were asked to empty their bladder immediately prior to this assessment.

@@

### Body Composition

Body composition, including body mass index (BMI), body fat percentage, lean muscle mass, body water, and bone mass, was assessed using multi-frequency bioelectrical impedance analysis (BIA). Measurements were performed using the Tanita BC-558 Iron-man Seg-mental Body Composition Monitor (Tanita Ironman, Ar-lington Heights, IL 60005 USA). Participants were instructed to fast for at least 4 hours prior to the assessment and to avoid strenuous exercise and alcohol consumption for at least 12 hours. They were also asked to empty their bladder before the measurement. Participants stood barefoot on the device's foot electrodes and held the hand electrodes while wearing light clothing.

@@

where $x_{i1},\dots,x_{i6}$ denote the six centered and scaled immune-cell predictors (absolute lymphocyte count, total B-cell count, and the four B-cell phenotypes by CD21/CD11c combinations). In the first model, we regressed $\widehat{\theta}_{ik}$ on those six immune predictors via @eq-second-model. In the second model, we extended @eq-second-model by adding four additional terms, $\beta_{k,\mathrm{age}} (\mathrm{age}_i)$, $\beta_{k,\mathrm{sex}} (\mathrm{sex}_i)$, $\beta_{k,\mathrm{fat}} (\mathrm{fat}_i)$ and $\beta_{k,\mathrm{muscle}} (\mathrm{muscle}_i)$, to account for confounding by age, sex and body composition variables when modeling RRi kinetics. Prior to fitting either multivariate model, all continuous predictors (immune counts and age) were centered to their mean and scaled by their standard deviation. We specified $\mathcal{N}(0,3)$ priors on every main-effect coefficient $\beta_{k}$ to exert a weakly regularizing influence and to mitigate the impact of outliers.

@@

Further model diagnostics, parameters and posterior statistics are available in the supplementary material.

@@

[**Table 1**]{#tab-1}. Sample characteristics in sociodemographical, body composition and immunological variables.

@@

data[, sex := factor(sex, levels = c("F", "M"), labels = c("Females", "Males"))]
tbl_data <- data[, c(15:21,9:14)]
tbl_data |> 
  tbl_summary(by = sex,
            missing = "no",
            statistic = list(
              all_continuous() ~ "{mean} ± {sd}"
            ),
            label = list(
              age ~ "Age (years old)",
              weight ~ "Weight (kg)",
              height ~ "Height (cm)",
              imc ~ "BMI (kg/m2)",
              muscle_total ~ "Muscle mass (kg)",
              fat_total ~ "Body fat (%)",
              abc_linfocitos_total ~ "Total lymphocytes (n)",
              abc_linfocitosb_total ~ "Total B cells (n)",
              cd21_m_cd11c_p ~ "CD21- CD11c+ B cells (n)",
              cd21_p_cd11c_p ~ "CD21+ CD11c+ B cells (n)",
              cd21_p_cd11c_m ~ "CD21+ CD11c- B cells (n)",
              cd21_m_cd11c_m ~ "CD21- CD11c- B cells (n)"
            ),
            digits = everything() ~ 2) |> 
  add_overall() |> 
  add_difference(test = list(all_continuous() ~ "smd")) |> 
  modify_spanning_header(all_stat_cols(stat_0 = F) ~ "**Sex**") |> 
  remove_abbreviation() |> 
  remove_footnote_header()
```

@@

![](../figures/hand-designed/rri-curve.png){width=100%}

@@

<!--- ## See output from "R/rri-characteristics.R"
   Parameters Estimate    SE               CI
       <fctr>    <num> <num>           <char>
1:      alpha   884.77 21.60 [840.27, 925.77]
2:       beta   403.99 24.57  [355.6, 451.58]
3:          c     0.89  0.02     [0.84, 0.94]
4:     lambda     3.01  0.22     [2.57, 3.44]
5:        phi     2.18  0.27     [1.63, 2.71]
6:        tau     6.89  0.11     [6.66, 7.11]
7:      delta     2.50  0.11     [2.28, 2.72]
--->

In our sample, the baseline RRi controlled by $\alpha$ was 884.77 ms (CI~95%~[840.27, 925.77]), with an exercise induced drop ($\beta$) of 403.99 (CI~95%~[355.6, 451.58]). The recovery proportion ($c$ parameter) relative to $\beta$ was 0.89 (CI~95%~[0.84, 0.94]).

In relation to the rate parameters, controlling the speed of transition between exercise and recovery phases, we observed a rate of exercise-induced decay ($\lambda$) of 3.01 min^-1^ (CI~95%~[2.57, 3.44]) and a post exercise recovery rate ($\phi$) of 2.18 min^-1^ (CI~95%~[1.63, 2.71]).

For time parameters controlling the timing of the aforementioned RRi kinetics, we observed that the time of the exercise-induced ($\tau$) drop aligned at 6.89 (CI~95%~[6.66, 7.11]) and a duration of the exertion-related RRi depression ($\delta$) was prolonged for 2.50 (CI~95%~[2.28, 2.72]).

@@

![](../figures/hand-designed/rri-signatures.png){width=100%}

@@

For individuals with more expression of both CD21 and CD11c (i.e., double positive), we observed a 99.5% posterior probability of higher RRi at rest (ES on $\alpha$ = 0.48, CI~95%~[0.11, 0.84], pd = 99.5%, ps = 97.9%), 97.3% posterior probability of more pronounced exercise-induced RRi decrease (ES on $\beta$ = 0.34, CI~95%~[0, 0.71], pd = 97.3%, ps = 91.5%), 91.5% posterior probability of less post-exercise RRi recovery proportion (ES on $c$ = -0.24, CI~95%~[-0.6, 0.1], pd = 91.5%, ps = 79.1%) and 90.5% posterior probability of higher duration of RRi depression till recovery kinetics begin, relative to the RRi drop (ES on $\delta$ = 0.26, CI~95%~[-0.12, 0.67], pd = 90.5%, ps = 79.5%).

@@

On the other side, individuals with higher counts of B cells absent of CD21 or CD11c markers (i.e., double negative) had 93.5% posterior probability of lower resting RRi (ES on $\alpha$ = -0.31, CI~95%~[-0.7, 0.09], pd = 93.5%, ps = 84.7%), 98.2% posterior probability of less exercise-induced RRi decrease (ES on $\beta$ = -0.41, CI~95%~[-0.8, -0.02], pd = 98.2%, ps = 94.2%), 96.1% posterior probability of better post-exercise recovery (ES on $c$ = 0.34, CI~95%~[-0.04, 0.72], pd = 96.1%, ps = 89.2%), 98% posterior probability of steeper RRi drop (ES on $\lambda$ = 0.44, CI~95%~[-0.01, 0.84], pd = 98%, ps = 94.4%) and 92.8% posterior probability of early initiation of cardiac autonomic kinematics (ES on $\tau$= -0.3, CI~95%~[-0.68, 0.14], pd = 92.8%, ps = 83.5%).

@@

For B cells with more CD11c expression, but not CD21, we observed that greater counts of this cell phenotype were associated 73% posterior probability of having a slower, or less steep, RRi drop initiated by exercise (ES on $\lambda$ = -0.24, CI~95%~[-0.61, 0.13], pd = 90.2%, ps = 76.8%) and 95.9% posterior probability of observing a delay in exercise-induced cardiac autonomic kinematics (ES on $\tau$ = 0.32, CI~95%~[-0.04, 0.68], pd = 95.9%, ps = 88.7%).

The posterior distribution for each effect for all B cell phenotypes on the different parameters controlling the exercise-induced RRi dynamics can be seen in [Figure 4](#fig-4).

![](../figures/fig_params_vs_immune_adjusted.png){width=100%}

[**Figure 4**](#fig-4). Posterior distribution of the standardized effect of B cell CD21 and CD11c expression on model parameters influencing the exercise-induced cardiac autonomic dynamics. $\alpha$ controls the baseline RRi; $\beta$ is proportional to the RRi drop, induced by exercise; $c$ is the recovery proportion, relative to $\beta$; $\lambda$ and $\phi$ are the steepness of exercise-induced RRi drop and post exercise recovery; $\tau$ correspond to the time at which the exercise-induced RRi drop occurs; $\delta$ correspond to the time-duration of the RRi depression, relative to $\tau$, that takes before the recovery kinetics begin.

@@

Our findings reveal a compelling dichotomy in the relationship between specific B cell subpopulations and autonomic dynamics during exercise in older adults. Individuals with a higher proportion of CD21^+^/CD11c^+^ (double positive) B cells exhibited characteristics of heightened parasympathetic activity at rest, including elevated baseline vagal tone and a longer resting RR intervals. This group also displayed an increased sympathetic response during exercise, indicated by an abrupt drop in RRi, a decreased cardiac recovery post-exercise and a longer time-to-recovery began. Conversely, participants with a predominance of CD21^-^/CD11c^-^ (double negative) B cells showed the opposite autonomic profile: low baseline vagal tone characterized by shorter RR and intervals. They also presented a decreased sympathetic response to physical exertion, reflected in a smaller, but steeper, drop in RRi, and a more close-to-baseline heart rate return following exercise. These contrasting observations suggest that the relative abundance of these B cell subpopulations may serve as an indicator of an individual's autonomic balance, with the CD21^+^/CD11c^+^ phenotype aligning with a vagally dominant state with increased sympathetic reactivity, while the CD21^-^/CD11c^-^ phenotype corresponds to a more sympathetically inclined state coupled with efficient compensatory vagal recovery capacity.

@@

The increased sympathetic response during exercise observed in this group is another notable finding. Moreover, a high resting vagal tone typically leads to a more pronounced vagal withdrawal at the onset of exercise, which aligns well with the observed data, suggesting a more pronounced initial heart rate acceleration, potentially due to increased vagal withdrawal and cholinergic modulation of the autonomic nervous system (ANS) [@giunta2024autonomic]. Cholinergic activation has been shown to counteract sympathetic signals, and a blunted adrenergic response might also limit the exercise-induced mobilization of lymphocytes [@smith2025exercise], potentially preventing marked immune activation in response to physical stress, thus suggesting a more controlled immune state in individuals with a dominant vagal influence.

However, this vagal-centric profile was also characterized by a slower return to baseline heart rate after exercise. Despite a less pronounced initial increase in heart rate, individuals with more CD21^+^/CD11c^+^ B cells took longer to recover at the same extent that individuals with greater CD21^-^/CD11c^-^. This protracted recovery could stem from a less efficient re-establishment of autonomic balance following a moderated sympathetic response, where the sympathetic nervous system (SNS) might not have been sufficiently activated to trigger a strong subsequent vagal rebound. Alternatively, it might indicate a degree of autonomic rigidity [@castillo2025enhancing], where a predominantly parasympathetic system at rest exhibits difficulty in rapidly adjusting after a perturbation. The lower $c$ parameter (recovery proportion of the RRi curve) observed in this group supports the notion of reduced autonomic flexibility, suggesting that while a basal anti-inflammatory environment might be present, the capacity for rapid adaptation to acute physiological challenges could be limited.

The characteristics associated with a higher proportion of CD21^+^/CD11c^+^ B cells find parallels in active older adults, who often display high resting HRV and moderated cardiovascular responses to exercise [@mabe2024associations], along with less inflammatory immune profiles [@tylutka2021lifestyle]. This suggests that these memory B cells might serve as both a marker and a modulator of this physiological state. While CD11c^+^ B cells in healthy individuals can differentiate into plasmablasts upon stimulation [@golinski2020cd11c], they also exhibit some resistance to regulatory signals like IL-10 [@golinski2020cd11c], implying potential modulation by neuroendocrine pathways. Future research into the effects of interventions such as chronic vagal nerve stimulation on these cells could be valuable in exploring ways to enhance their anti-inflammatory profile while preserving their capacity for immune response when needed.

@@

During exercise, this subpopulation exhibited an amplified autonomic response, with the marked tachycardia, reflected in a lower $\alpha$ and greater $\lambda$ parameters, indicating a robust sympathetic discharge. This heightened response is consistent with their low baseline parasympathetic tone; with less vagal "braking", the cardiovascular system responds rapidly and intensely to the exercise stimulus. While potent acute sympathetic activation has advantages such as increased cardiac output and immune cell mobilization [@smith2025exercise], potentially contributing to the clearance of senescent cells through exercise [@simpson2021exercise], it also carries potential risks, including cardiac arrhythmias and increased inflammation.

A notable and potentially beneficial finding was the better autonomic recovery (high $c$ parameter) observed in the CD21^-^/CD11c^-^ group despite their lower initial vagal tone. This suggests a compensatory and efficient vagal reactivation following the sympathetic surge, possibly mediated by baroreceptor and mechanoreceptor reflexes responding to the exercise-induced cardiovascular changes [@besnier2017exercise]. This efficient cholinergic recovery could aid in resolving immune activation once the exercise stimulus ceases. Thus, individuals with a more sympathicotonic profile might be capable of mounting strong but transient responses, followed by a relatively rapid return to autonomic balance mediated by the parasympathetic nervous system. This pattern resonates with descriptions of mild autonomic insufficiency, where sympathetic hyper-responsiveness can coexist with late vagal hyper-reactivity [@besnier2017exercise], and aligns with the concept of autonomic resilience [@castillo2025enhancing].

This autonomic and immune profile shares similarities with clinical conditions such as metabolic syndrome and obesity, where reduced HRV and high basal sympathetic activity are often observed, although some form of interventions can improve vagal reactivity [@tylutka2021lifestyle; @frasca2023obesity]. The link between alterations in B cells and autonomic behavior suggests a bidirectional relationship, where not only can the immune profile influence the autonomic response, but also that modulating the ANS can potentially impact age-associated immune alterations, as seen with chronic exercise interventions [@weyh2020physical; @tylutka2021lifestyle].

@@

This study offers a novel contribution by integrating autonomic physiology, through nonlinear modeling of RRi dynamics, with detailed B-cell phenotyping in the context of healthy aging. The application of a coupled logistic framework allowed us to capture temporally resolved autonomic shifts during exercise and recovery, providing a more dynamic assessment compared to standard HRV metrics, while controlling for key demographical and body composition variables. Our findings, which highlight distinct associations between CD21/CD11c B-cell subsets and specific autonomic parameters, offer new avenues for understanding immune-ANS interactions.

However, several limitations warrant consideration. The cross-sectional design of our study precludes the establishment of causal inferences between the observed immune and autonomic characteristics. Furthermore, while surface marker-based phenotyping allowed for the identification of specific B-cell subpopulations, it did not elucidate their functional or mechanistic relevance. The absence of functional assays or transcriptomic profiling to complement the flow cytometry data, along with the unquantified potential for batch or gating effects, represents a further limitation. Methodologically, our two-stage modeling pipeline did not propagate posterior uncertainty between steps, underscoring the need for future work employing fully joint hierarchical inference.

Moving forward, future research should aim to overcome these limitations by employing longitudinal study designs to explore causality, and incorporating functional and mechanistic analyses to understand the roles of the identified B-cell populations to further elucidate the complex interplay between the immune and autonomic systems in aging.

@@

This research suggests that B cell subpopulations reflect and potentially modulate ANS dynamics during exercise in older adults. An integrative neuro-immunological approach provides a more comprehensive understanding of the health status of older individuals, identifying "pro-resilient" versus "pro-fragile" profiles that might not be evident when evaluating systems separately. Individuals with a predominance of CD21^+^/CD11c^+^ B cells exhibit higher baseline vagal tone, an increased sympathetic response during exercise, but an incomplete autonomic recovery. In contrast, those with CD21-/CD11c- B cells present lower baseline vagal tone, a more pronounced sympathetic response, and more complete post-exercise recovery. These results indicate that CD21^+^/CD11c^+^ B cells are associated with a more regulated and less inflammatory autonomic profile, whereas CD21^-^/CD11c^-^ B cells reflect a state of chronic inflammation. Exercise could modulate these autonomic and immune profiles, thereby improving cardiovascular health in older adults. This investigation underscores the importance of an integrative neuro-immunological approach for developing predictive biomarkers and personalized strategies to optimize health in aging.
